---
title: "2020 Police Employee Data Import"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

The goal of this notebook is to build a categorized list of police agencies in Texas. See the README for information about where the data comes from.

In this notebook we will:

- Import the four files
- Clean if necessary
- Add a category of city, county, school, state, etc
- Merge the files together

In another notebook we might do further categorization to note agencies from Central Texas if we can find a programattic way to do that.

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(readxl)
library(janitor)
library(DT)
```

We deal with each data file because they have different import strategies

## City agencies

Import the data

```{r city-import}
city <- read_excel("data-raw/Table_78_Full_time_Law_Enforcement_Employees_Texas_by_Cities_2020.xlsx", skip = 5) %>% clean_names()

city %>% glimpse()
```

### Clean cities

- add a type
- rename columns

```{r city-clean}
city_cleaned <- city %>% 
  mutate(type = "City") %>% 
  select(
    type,
    agency = city,
    everything()
  )

city_cleaned %>% head()
```

### Pare city

```{r city-pare}
city_pared <- city_cleaned %>% 
  select(type, agency, total_officers)

city_pared %>% glimpse()
```


## County agencies

```{r county-import}
county <- read_excel("data-raw/Table_80_Full_time_Law_Enforcement_Employees_Texas_by_Metropolitan_and_Nonmetropolitan_Counties_2020.xlsx", skip = 5) %>% clean_names()

county %>% head
```

### Clean counties

- Fill down the metro/non-metro columns
- create type column
- rename for consistency

```{r county-clean}
county_cleaned <- county %>%
  fill(metropolitan_nonmetropolitan) %>%
  mutate(
    metropolitan_nonmetropolitan = str_remove(metropolitan_nonmetropolitan, " Counties"),
    type = "County"
  ) %>% 
  select(
    type,
    metro = metropolitan_nonmetropolitan,
    agency = county,
    everything()
  )

county_cleaned %>% head()
```

### Pare counties

```{r county-pare}
county_pared <- county_cleaned %>% 
  select(type, agency, total_officers)

county_pared %>% glimpse()
```


## Universities and Colleges

Two notes from the file:

1 The student enrollment figures provided by the United States Department of Education are for the 2019 school year, the most recent available. The enrollment figures include full-time and part-time students.
2 Student enrollment figures were not available. (values are blank - crit)

```{r school-import}
school <- read_excel(
    "data-raw/Table_79_Full_time_Law_Enforcement_Employees_Texas_by_University_and_College_2020.xlsx",
    skip = 5,
    n_max = 69
  ) %>% clean_names()

school
```

### Schools clean

- remove some note numbers from school and campus columns
- add the type
- reorder/rename columns

```{r school-clean}
school_cleaned <- school %>% fill(universities_college) %>% 
  mutate(
    campus = str_remove(campus, "\\d"),
    universities_college = str_remove(universities_college, "[:\\d]") %>% trimws(),
    type = "School"
  ) %>% 
  select(
    type,
    agency = universities_college,
    description = campus,
    student_enrollment = student_enrollment1,
    everything()
  )

school_cleaned
```

### Pare schools

```{r school-pare}
school_pared <- school_cleaned %>% 
  select(type, agency, description, total_officers)

school_pared %>% glimpse()
```


## State agencies

```{r state-import}
state <- read_excel("data-raw/Table_81_Full_time_Law_Enforcement_Employees_Texas_by_State_Tribal_and_Other_Agencies_2020.xlsx", skip = 4) %>% clean_names()

state %>% head(10)
```

### Clean state agencies

- Fill columns
- remove `:` from compound names
- rename/reorder columns


```{r state-clean}
state_cleaned <- state %>% 
  fill(state_tribal_other, agency) %>% 
  mutate(
    agency = str_remove(agency, ":") %>% trimws()
  ) %>% 
  select(
    type = state_tribal_other,
    agency,
    description = unit_office,
    everything()
  )

state_cleaned %>% head()
```

### Pare state agencies

```{r state-pare}
state_pared <- state_cleaned %>% 
  select(type, agency, description, total_officers)

state_pared %>% glimpse()
```

## Combining

```{r combine}
agencies_combined <- school_pared %>% 
  bind_rows(state_pared) %>% 
  bind_rows(county_pared) %>% 
  bind_rows(city_pared)

agencies_combined %>% glimpse()
```


## Searchable table

```{r}
agencies_combined %>% 
  datatable()
```



## Export

```{r export}
agencies_combined %>% write_rds("data-processed/ped_texas_2020.rds")
agencies_combined %>% write_csv("data-processed/ped_texas_2020.csv")
```

