---
title: "Texas Police Agencies 2018"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

This notebook explores and ranks Texas police agencies according to Police Employee Data from the [FBI Crime Data Explorer](https://crime-data-explorer.fr.cloud.gov/downloads-and-docs). See the [README.md](https://github.com/utdata/rwd-police-agencies) for more information.

## Dependencies

I had trouble importing the original data into RStudeio, so you have to run the "01-file-process.ipyng" notebook first, which needs csvkit and jupyter notebooks.


```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
```

## Import data

There are some duplicate rows, so we include the `unique()` function.

```{r}
tx_2018 <- read_csv("data/pe_texas_2018.csv") %>% unique()
```


## Explore

Peek at the data

```{r}
tx_2018 %>% head()
```

Get the column names

```{r}
tx_2018 %>% names()
```

## Types of agencies

```{r}
tx_2018 %>% 
  count(agency_type_name)
```

## Rank all agencies by officer count

```{r}
tx_2018 %>% 
  arrange(officer_ct %>% desc()) %>% 
  select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(50)
```

## filter to just city & county types

```{r}
tx_local_2018 <- tx_2018 %>% 
  filter(agency_type_name %in% c("City","County")) %>% 
  arrange(officer_ct %>% desc()) %>% 
  select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct)

tx_local_2018
tx_local_2018 %>% write_csv("data/tx_local_2018.csv")
```



## Explore Independent School District:

```{r}
tx_ed_2018 <- tx_2018 %>% 
  filter(pub_agency_name == "Independent School District:", officer_ct > 50) %>% 
  arrange(officer_ct %>% desc()) %>% 
  select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(10)

tx_ed_2018
```

What is the United school district?

```{r}
tx_2018 %>% 
  filter(pub_agency_name == "Independent School District:",  pub_agency_unit == "United") %>% 
  arrange(officer_ct %>% desc()) %>% 
  # select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head()

```


## Explore University police

```{r}
tx_uni_2018 <- tx_2018 %>% 
  filter(agency_type_name == "University or College") %>% 
  arrange(officer_ct %>% desc()) %>% 
    select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(10)

tx_uni_2018
```

## Austin MSA agencies

Putting together a list of Austin MSA agencies for another project, utdata/rwd-r-leso. 

```{r centex}
centex_name <- c(
  "BASTROP",
  "CALDWELL",
  "HAYS",
  "TRAVIS",
  "WILLIAMSON"
)

# study to combine name and unit
# needs removal of NA
tx_2018 %>% 
  mutate(
    agency = paste(pub_agency_name, pub_agency_unit)
  ) %>%
  # select(agency) %>% # remove late
  mutate(
    agency = str_remove(agency, " NA")
  )

# create list of non-state agencies in ausitn msa
austin_msa <- tx_2018 %>% 
  filter(county_name %in% centex_name) %>% 
  select(
    pub_agency_name,
    pub_agency_unit,
    county_name,
    agency_type_name
  ) %>% 
  # fix agency name
  mutate(
    agency = paste(pub_agency_name, pub_agency_unit),
    agency = str_remove(agency, " NA")
  ) %>%
  select(agency, county_name, agency_type_name) %>% 
  filter(
    agency_type_name != "Other State Agency"
  ) %>% 
  arrange(county_name, agency_type_name, agency)

## list it
austin_msa

```



## Output files for PIRs

We're going to look at several kinds of agencies:

- Top 10 Universities and Colleges by officer count.
- City and county together based on officer count to fill out the rest in the class.
- a list of Austin MSA agencies for another project

```{r}
tx_pir_police_agencies <- bind_rows(
  tx_uni_2018,
  tx_local_2018
)

tx_pir_police_agencies %>% 
  write_csv("data/tx_pir_police_agencies.csv")

austin_msa %>% 
  write_csv("data/austin_msa_agencies.csv")
```

Of note, I'm not using school districts in the above list,
