---
title: "Texas Police Agencies 2018"
output:
  html_document:
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

This notebook explores and ranks Texas police agencies according to data from the [FBI Crime Data Explorer](https://crime-data-explorer.fr.cloud.gov/downloads-and-docs). See the [README.md](README.md) for more information.

## Dependencies

I had trouble importing the original data into RStudeio, so you have to run the "01-file-process.ipyng" notebook first, which needs csvkit and jupyter notebooks.


```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
```

## Import data

There are some duplicate rows, so we include the `unique()` function.

```{r}
tx_2018 <- read_csv("data/pe_texas_2018.csv") %>% unique()
```


## Explore

Peek at the data

```{r}
tx_2018 %>% head()
```

Get the column names

```{r}
tx_2018 %>% names()
```

## Types of agencies

```{r}
tx_2018 %>% 
  count(agency_type_name)
```

## Rank all agencies by officer count

```{r}
tx_2018 %>% 
  arrange(officer_ct %>% desc()) %>% 
  select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(50)
```

## filter to just city & county types

```{r}
tx_local_2018 <- tx_2018 %>% 
  filter(agency_type_name %in% c("City","County")) %>% 
  arrange(officer_ct %>% desc()) %>% 
  select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(30)

tx_local_2018
```



## Explore Independent School District:

```{r}
tx_ed_2018 <- tx_2018 %>% 
  filter(pub_agency_name == "Independent School District:", officer_ct > 50) %>% 
  arrange(officer_ct %>% desc()) %>% 
  select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(10)

tx_ed_2018
```

What is the United school district?

```{r}
tx_2018 %>% 
  filter(pub_agency_name == "Independent School District:",  pub_agency_unit == "United") %>% 
  arrange(officer_ct %>% desc()) %>% 
  # select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head()

```


## Explore University police

```{r}
tx_uni_2018 <- tx_2018 %>% 
  filter(agency_type_name == "University or College") %>% 
  arrange(officer_ct %>% desc()) %>% 
    select(pub_agency_name, agency_type_name, pub_agency_unit, officer_ct) %>% 
  head(10)

tx_uni_2018
```

## Output files for PIRs

We're going to look at several kinds of agencies:

- Top 10 Universities and Colleges
- School districts with more than 50 officers
- City and county together based on officer count to fill out the rest in the class

```{r}
tx_pir_police_agencies <- bind_rows(
  tx_uni_2018,
  tx_ed_2018,
  tx_local_2018
)

tx_pir_police_agencies %>% 
  write_csv("data/tx_pir_police_agencies.csv")
```

